---
title: "Homework 1"
author: "Megan Marziali"
date: "September 13, 2020"
output: github_document
---

This document contains a solution for Homework 1 for the Building Blocks section of P8105.

## Problem 1

### Creation of data frame

The code chunk below involves the creation of a data frame with various types of variables. 

```{r}
library(tidyverse)

prob_one_df = tibble(
  vec_norm = rnorm(10),
  vec_char = c("the", "quick", "brown", "fox", "jumped", "over", "the", "lazy", "dog", "today"),
  vec_factor = factor(c("the", "the", "the", "fuzzy", "fuzzy", "fuzzy", "cat", "cat", "cat", "cat")),
  # Denotes whether values in this data frame are greater than 0 or not
  vec_logical = vec_norm > 0,
)
```

### Calculation of means

The means for each variable in the data frame are calculated below.

```{r}
mean(pull(prob_one_df, vec_norm))
mean(pull(prob_one_df, vec_char))
mean(pull(prob_one_df, vec_factor))
mean(pull(prob_one_df, vec_logical))
```

The mean cannot be computed for the character and factor vectors, as they do not take on numerical values, which explains the occurrence of the above warnings. The mean was computed for the logical variable, although it likely should not have been computed as it relates instead to the number of "true" responses.

### Variable conversions

This code chunk converts variables from one type to another.

```{r, echo=TRUE}
as.numeric(pull(prob_one_df, vec_logical))
as.numeric(pull(prob_one_df, vec_char))
as.numeric(pull(prob_one_df, vec_factor))
```

The above conversions explain some of the findings from the mean calculations. Firstly, the numeric conversion demonstrates that R is reading the logical vector as a binary (0, 1) which explains why the mean was computed. The character variable cannot be converted to numeric. When converting the factor to numeric, it is converted as the three levels of the variable.

The next code chunk does additional variable conversions.

```{r}
as.numeric(pull(prob_one_df, vec_logical)) * (pull(prob_one_df, vec_norm))
as.factor(pull(prob_one_df, vec_logical)) * (pull(prob_one_df, vec_norm))

fact_of_log = as.factor(pull(prob_one_df, vec_logical))
fact_num_by_random = as.numeric(fact_of_log) * (pull(prob_one_df, vec_norm))
fact_num_by_random
```

The warning above occurring for the fact_log_random variable is due to an attempt to multiply a non-numeric variable by a numeric variable.

## Problem 2

### Descriptive Statistics

The code chunk below is designed to call the correct package.
```{r}
data("penguins", package = "palmerpenguins")
```

The penguins data set includes variables `r names(penguins)`. It has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The **species** variable is a `r class(penguins$"species")` variable. This factor variable includes levels `r levels(penguins$"species")`: `r sum(penguins$"species" == "Adelie")` observations are classified as Adelie, `r sum(penguins$"species" == "Chinstrap")` as Chinstrap and `r sum(penguins$"species" == "Gentoo")` as Gentoo. There are `r sum(is.na(penguins$"species"))` missing values for this variable.

The **island** variable is a `r class(penguins$"island")` variable. This factor variable includes levels `r levels(penguins$"island")`: `r sum(penguins$"island" == "Biscoe")` observations are classified as Biscoe, `r sum(penguins$"island" == "Dream")` as Dream and `r sum(penguins$"island" == "Torgersen")` as Torgersen. There are `r sum(is.na(penguins$"island"))` missing values for this variable.

The **bill length** variable is a `r class(penguins$"bill_length_mm")` variable. The mean bill length is `r mean(penguins$"bill_length_mm", na.rm = TRUE)` with a standard deviation of `r sd(penguins$"bill_length_mm", na.rm = TRUE)`; the median is `r median(penguins$"bill_length_mm", na.rm = TRUE)` and interquartile (IQR) range of `r IQR(penguins$"bill_length_mm", na.rm = TRUE)`. There are `r sum(is.na(penguins$"bill_length_mm"))` missing values for this variable.

The **bill depth** variable is a `r class(penguins$"bill_depth_mm")` variable. The mean bill length is `r mean(penguins$"bill_depth_mm", na.rm = TRUE)` with a standard deviation of `r sd(penguins$"bill_depth_mm", na.rm = TRUE)`; the median is `r median(penguins$"bill_depth_mm", na.rm = TRUE)` and interquartile (IQR) range of `r IQR(penguins$"bill_depth_mm", na.rm = TRUE)`. There are `r sum(is.na(penguins$"bill_depth_mm"))` missing values for this variable.

The **flipper length** variable is a `r class(penguins$"flipper_length_mm")` variable. The mean bill length is `r mean(penguins$"flipper_length_mm", na.rm = TRUE)` with a standard deviation of `r sd(penguins$"flipper_length_mm", na.rm = TRUE)`; the median is `r median(penguins$"flipper_length_mm", na.rm = TRUE)` and interquartile (IQR) range of `r IQR(penguins$"flipper_length_mm", na.rm = TRUE)`. There are `r sum(is.na(penguins$"flipper_length_mm"))` missing values for this variable.

The **body mass** variable is a `r class(penguins$"body_mass_g")` variable. The mean bill length is `r mean(penguins$"body_mass_g", na.rm = TRUE)` with a standard deviation of `r sd(penguins$"body_mass_g", na.rm = TRUE)`; the median is `r median(penguins$"body_mass_g", na.rm = TRUE)` and interquartile (IQR) range of `r IQR(penguins$"body_mass_g", na.rm = TRUE)`. There are `r sum(is.na(penguins$"body_mass_g"))` missing values for this variable.

The **sex** variable is a `r class(penguins$"sex")` variable. This factor variable includes levels `r levels(penguins$"sex")`: `r sum(penguins$"sex" == "Male")` observations are classified as male and `r sum(penguins$"sex" == "Female")` as female. The number of missing observations for this variable is `r sum(is.na(penguins$"sex"))`. There are `r sum(is.na(penguins$"sex"))` missing values for this variable.

The **year** variable is a `r class(penguins$"year")` variable, with `r length(penguins$"year")` observations. The year variable has a minimum of `r min(penguins$"year")` and a maximum of `r max(penguins$"year")`. There are `r sum(is.na(penguins$"year"))` missing values for this variable.

### Plot

The code chunk below creates a plot of bill length (x) and flipper length (y). Species are denoted with different colors.

```{r, warning = FALSE}
# The library call below is purely for aesthetic reasons, to be able to alter the colors in the plot.

library(wesanderson)

fliper_length_plot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

fliper_length_plot + scale_color_manual(values = wes_palette( n = 3, name = "Moonrise2"))

# The code below is to save the scatterplot.

ggsave("flipper_length_plot.png")
```




